---
title: "Lab 1"
author: "Tyron Samaroo"
output: pdf_document
date: "11:59PM February 8, 2020"
---

You should have RStudio installed to edit this file. You will write code in places marked "TO-DO" to complete the problems. Some of this will be a pure programming assignment. The tools for the solutions to these problems can be found in the class practice lectures. I want you to use the methods I taught you, not for you to google and come up with whatever works. You won't learn that way.

To "hand in" the homework, you should compile or publish this file into a PDF that includes output of your code. Once it's done, push by the deadline to your repository in a directory called "labs".

* Print out the numerical constant pi with ten digits after the decimal point using the internal constant `pi`.

```{r}
options(digits = 10)
pi
```


* Sum up the first 100 terms of the series 1 + 1/2 + 1/4 + 1/8 + ...

```{r}
1/2 ^ (0:99) #Generates 1 + 1/2 ...
sum(1/2 ^ (0:99)) 
```

* Find the product of the first 20 terms of `1/3 * 1/6 * 1/9 *` ...

```{r}
#
prod(1 / seq(3, 60, by = 3))

```


* Find the product of the first 500 terms of `1 * 1/2 * 1/4 * 1/8 *` ...

```{r}
prod(1/2 ^ (0:499))
```

Is this answer *exactly* correct? 

No because will give wrong computation. Not precise answer. The answer is almost 0 but not exactly 0

* Figure out a means to express the answer more exactly. Not compute exactly, but express more exactly.

```{r}
#Use Ln to help.
log(1/2 ^ (0:499))
sum(log(1/2 ^ (0:499)))

```

* Create the sequence `x = [Inf, 20, 18, ..., -20]`.

```{r}
#Use c to concat
x = c(Inf,seq(from = 20, to = -20, by = -2))
x

```

Create the sequence `x = [log_3(Inf), log_3(100), log_3(98), ... log_3(-20)]`.

```{r}
x = c(Inf,seq(from = 100, to=-20, by = -2))
# Getting NaN no good
x = log(x, base = 3)
x
```

Comment on the appropriateness of the non-numeric values.

Nan come from log(-num). Inf come from log3(Inf and log -inf come from log3(0)


* Create a vector of booleans where the entry is true if `x[i]` is positive and finite.

```{r}
x
x > 0 & x != Inf & !is.nan(x)
is_pos_real_bool = x > 0 & x != Inf & !is.nan(x)
is.finite(x) #Another way
is_pos_real_bool


```

* Locate the indices of the non-numbers in this vector. Hint: use the `which` function.

```{r}
#Theres are values that arent finite 
which(!is_pos_real_bool) 

```

* Locate the indices of the infinite quantities in this vector. Hint: use the `which` function.

```{r}

x
which(x == Inf | x == -Inf)
is.infinite(x) # Return vector location of -inf and inf
which(is.infinite(x)) # return index of those location



```

* Locate the indices of the min and max in this vector. Hint: use the `which.min` and `which.max` functions.

```{r}
x[is.infinite(x)] = NA 
x
which.min(x)
which.max(x)
```

* Count the number of unique values in `x`.

```{r}
#TO-DO
x
length(unique(x))
```

* Cast `x` to a factor. Do the number of levels make sense?

```{r}
factor(x)
#as.factor(x) # Same
```

* Cast `x` to integers. What do we learn about R's infinity representation in the integer data type?

```{r}
x
as.integer(x)
```

* Use `x` to create a new vector `y` containing only real numbers.

```{r}
x
y = x[is.finite(x)]
y
length(y) #Different size of vector
```

* Use the left rectangle method to numerically integrate x^2 from 0 to 1 with rectangle size 1e-6.

```{r}
sum(seq(0, 1 - 1e-6, by = 1e-6) ^ 2) * 1e-6
```


* Calculate the average of 100 realizations of standard Bernoullis in one line using the `sample` function.

```{r}
mean(sample(c(0,1) ,size = 100, replace = TRUE))
```


* Calculate the average of 500 realizations of Bernoullis with p = 0.9 in one line using the `sample` function.

```{r}
#TO-DO
#Probability of 1 is 0.9
# realization_bernoullis = sample(c(0, 1),
#                                 size = 500,
#                                 replace = TRUE,
#                                 prob = c(0.1, 0.9))
# mean(realization_bernoullis)


mean(sample(c(0,1),size = 500, replace = TRUE, prob = c(0.1, 0.9)))
```

* In class we considered a variable `x_3` which measured "criminality". We imagined L = 4 levels "none", "infraction", "misdimeanor" and "felony". Create a variable `x3` here with 100 random elements (equally probable). Create it as a nominal (i.e. unordered) factor.

```{r}
x_3 = sample(c("none", "infraction", "misdimeanor", "felony"), 
             size = 100, 
             replace = TRUE, 
             prob = c(.25, .25, .25, .25))
factor(x_3)
```

* Use `x_3` to create `x_3_bin`, a binary feature where 0 is no crime and 1 is any crime.

```{r}
x_3
x_3_bin = ifelse(x_3 == "none", 1, 0)
x_3_bin
```

* Use `x_3` to create `x_3_ord`, an ordered, nominal factor variable. Ensure the proper ordinal ordering.

```{r}
sample_criminality= sample(c("none", "infraction", "misdimeanor", "felony"), 
             size = 100, 
             replace = TRUE)
x_3_ord = factor(sample_criminality, levels = c("none", "infraction", "misdimeanor", "felony"), ordered = TRUE)
x_3_ord
```
